Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-26 Manuscript/manuscript.Rmd
edits

After obtaining our optimally built random forest classifier, we aimed to leverage the ChIP-seq peak regions defined by the SMC3, ZNF143, CTCF, and RAD21 tanscription factors for precised TAD bounday prediction at base pair resolution. We found that when we validated our model against chromosome specific base pairs, we were able to identify clusters of regions where the model predicted the base pair to be a TAD boundary with probability equal to 1. By clustering the regions and performing medoid partitioning (See Methods), we were able to identify a comparative number of TAD boundaries as the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 10). To compare between predicted and called TAD boundaries we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flank of each. We found that the peak signal strength for the SMC3, ZNF143, CTCF, and RAD21 TFBS were much closer, on average, to our predicted TAD boudaries the the called TAD boundaries obtained using the ARROWHEAD pipeline for 10 kb resolution Hi-C data (Figure 5; Supplementary Figure ). ChIP-seq assays offer much higher resolutions (often in the tens of bases) compared to Hi-C sequenced data, which can be in the hundreds of kilobases. This disparity in resolution can affect a TAD caller's ability to precisely locate TADs. Moreover, by building a model using annotations known to be enriched at TAD boundaries we were able to bridge this gap in resolution and more precisely locate TAD boundaries that corresponded to the highest concentration in signal strength for top ranked TFBS.

>>>>

After obtaining our optimally built random forest classifier, we aimed to leverage the ChIP-seq peak regions defined by the SMC3, ZNF143, CTCF, and RAD21 tanscription factors for precise TAD bounday prediction at base pair resolution. We found that when we validated our model against chromosome specific base pairs, we were able to identify clusters of regions where the model predicted the base pair to be a TAD boundary with probability equal to 1. By clustering the regions and performing medoid partitioning (See Methods), we were able to predict a comparative number of TAD boundaries (base pairs) as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 10). To compare between predicted and called TAD boundaries we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flank of each. We found that the peak signal strength for the SMC3, ZNF143, CTCF, and RAD21 TFBS were much closer, on average, to our predicted TAD boudaries the the called TAD boundaries obtained using the ARROWHEAD pipeline for 10 kb resolution Hi-C data (Figure 5; Supplementary Figure ). ChIP-seq assays offer much higher resolutions (often in the tens of bases) compared to Hi-C sequenced data, which can be in the hundreds of kilobases. This disparity in resolution can affect a TAD caller's ability to precisely locate TADs. Moreover, by building a model using annotations known to be enriched at TAD boundaries we were able to bridge this gap in resolution and more precisely locate TAD boundaries that corresponded to the highest concentration in signal strength for top ranked TFBS.
